### 2016-09-28 paper.tex
changes

\newcommand{\zero}{\text{\textbf{0}}}
\newcommand{\Zproj}{Z^{\textit{proj}}}
\newcommand{\Zreopt}{Z^{\textit{owa}}}
\newcommand{\nreopt}{n^{\textit{owa}}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\matW}{\hat W}
\newcommand{\matV}{\hat V}
\newcommand{\W}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Wowa}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Waff}{\mathcal{\hat A}}
\newcommand{\WaffE}{{\mathcal{\hat A}_\E}}
\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\ahat}{\hat\alpha}
\newcommand{\afull}{\ahat^{\textit{full}}}
\newcommand{\wowa}{\hat\w^{owa}}
\newcommand{\wowafull}{\hat\w^{\textit{owa,full}}}
\newcommand{\wowastar}{\hat\w^{\textit{owa,*}}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\wtave}{\E\hat\w^{ave}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

@@

@@

%That is, we assume there exists a constant $c$ such that
%\begin{equation}
%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{\frac{d}{n}}} \le \exp\left(-ct^2\right)
%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{d/n}} \le \exp\left(-ct^2\right)
%%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(-\sigma^2t^2/2\right)
%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(\frac{-\sigma^2t^2}{2}\right)
%.
%\end{equation}
\begin{assumption}
Let $t>0$ and $\sigma^2$ be a constant called the \emph{variance proxy}.
Then with probability at least $1-\exp(-\sigma^2t^2/2)$,
\begin{equation}
%\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < \frac{t}{\sqrt n}
,
\end{equation}
where $\I_\wstar$ is the Fisher information matrix.
\end{assumption}

This is a mild condition that is known to hold in many situations of interest.
In the asymptotic regime when $n\to\infty$,
%this condition is known as root-$n$ consistency.
%In this case,
the variance proxy $\sigma^2=1$,
and the distribution of the estimation error is truly Gaussian.
Theorem 7.5.2 of \cite{lehmann1999elements} provides a particularly simple exposition.
Lehman's theorem requires only that $f$ be three times differentiable,
the data points be i.i.d.,
and some mild identifiability conditions.
Very strong results are known in the asymptotic regime.
For example, Condition 1 is satisfied even without the i.i.d. assumption.
See for example \cite{spokoiny2012parametricestimation}.

Similar results hold in the non-asymptotic case $n<\infty$,
but known results are less strong.
Most non-asymptotic results of this form require that the data points be sub-Gaussian.
For example, \cite{negahban2009unified} considers the case when the data points are sub-Gaussian, the likelihood satisfies a ``restricted strong convexity condition,'' and the regularizer is decomposable.
More recently, \cite{sivakumar2015beyond} showed Condition 1 holds when the data are only sub-exponential.
The strongest non-asymptotic results on the estimation error known to the authors is due to \cite{spokoiny2012parametricestimation}.
Spokoiny requires only a mild exponential moment condition (and does not even require the data be i.i.d.);
but can show that Condition 1 is satisfied only for $t < O(n)$.

Our condition is more general than the conditions used in previous work.
\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded,
and every bounded random variable is automatically sub-Gaussian by definition.

In conclusion,
there are many ways to verify that Condition 1 holds.
Pick whichever one suits your problem.
With this condition assumed, our results are elementary.

@@

\begin{theorem}
\label{thm:wave}
Assume Condition 1 holds.
Let $t>0$.
Then with probability at least $1-\exp(-t)$,
\begin{equation}
\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wmle_i} + \sqrt\frac{v_t}{mn}
\end{equation}
where
\begin{equation}
v_t =
\sigma^2
\left(
\tr{\I^{-1}_{\wstar}}
+ 2\sqrt{\tr \left({\I^{-2}_{\wstar}}\right)t}
+ 2\ltwo{\I^{-1}_{\wstar}}t
\right)
\end{equation}
\end{theorem}

@@

Now we consider the variance.
We have that
\begin{align}
\ltwobig{\wave-\E\wave}
&=
%\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave
%\\&=
%\ltwobig{\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave}
%\label{eq:var1}
%\\&=
\ltwobig{\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)}
\\&=
\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
\\&\sim
\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\subnormal{\sigma^2}}
\\&\sim
\frac{1}{\sqrt m}\ltwobig{\I^{-1/2}_\wstar\subnormal{\sigma^2}}
%\\&=
%\frac{1}{\sqrt {nm}}\normal\zero{\I^{-1}_{\E\wave}}
\label{eq:var2}
%\\&\sim
%\frac{1}{m}\normal{m\E\wmle_i}{\frac{m}{n}I^{-1}}
%\\&\sim
%\normal{\E\wmle_i}{\frac{1}{nm}I^{-1}}
%\\
%\wave-\E\wmle_i
%&\sim
%\normal{\zero}{\frac{1}{nm}I^{-1}}
%\\
%\wave-\E\wave
%&\sim
\end{align}
%Centering the distribution and applying Equation \ref{eq:expwave} gives
%\begin{equation}
%\E\wave-\wave
%\sim
%\normal{\zero}{\frac{1}{nm}I^{-1}}
%\end{equation}
In the last step we used the fact that adding a centered Gaussian to itself $m$ times is equivalent to multiplying the Gaussian by $\sqrt{m}$.
%So bounding the variance reduces to bounding the norm of a Gaussian.
Our last step is to bound the norm of the Gaussian in Equation \ref{eq:var2}.
Applying the bound in Proposition 1.1 from \cite{hsu2012tail} gives that with probability at least $1-\exp(-t)$,
\begin{equation}
\ltwo{\E\wave-\wave} \le \sqrt\frac{v_t}{nm}
\label{eq:ct1}
\end{equation}
%where
%\begin{equation}
%\begin{aligned}
%v_t'
   %= \tr\left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)
%+ 2&\sqrt{\tr \left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)^2t}&
%\\+ 2\ltwo{\frac{1}{nm}\I^{-1}_{\wtave}}t&
%\end{aligned}
%\end{equation}
%Then by the linearity of trace and scalability of norms,
%\begin{equation}
%v_t' =
%\frac{1}{mn}\left(
%\tr\I^{-1}_{\wtave}
%+ 2\sqrt{\tr \I^{-2}_{\wtave}t}
%+ 2\ltwo{\I^{-1}_{\wtave}}t
%\right)
%\label{eq:ct2}
%\end{equation}
Substituting Equations \ref{eq:expwave} and \ref{eq:ct1} into Equation \ref{eq:biasvar} gives the stated result.
\end{proof}

Note that this reduction in variance is essentially optimal.
By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
we can show that the variance of the maximum likelihood estimator trained on all of the data is
\begin{equation}
\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
\label{eq:var3}
\end{equation}
The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.

@@

\begin{theorem}
Assume Equation \ref{eq:assumption} holds.
Let $t>0$.
Then we have that with probability at least $(1-\exp(-t))^2$,
\begin{equation}
\ltwo{\wstar-\proj\W\wstar}
\le
%b_t\sqrt{\left(1-\frac{m}{d}\right)}
b_t\sqrt{\left(1-\frac{m}{d}\right)}
+
\sqrt\frac{v_{t}}{nm}
\end{equation}
where
%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
\begin{equation}
b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
\end{equation}
%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
and $v_t$ is defined as in Theorem \ref{thm:wave}.
\end{theorem}

In the remainder of this section,
we first state some simple conditions wherein the central limit theorem holds;
then we prove Theorems 1 and 2.

\subsection { The Central Limit Theorem}
\label{sec:clt}

The assumption in Equation \ref{eq:assumption} is motivated by the central limit theorem for maximum likelihood estimation.
There are many versions of this theorem.
Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
The notation has been changed to match ours.
\begin{theorem}
\label{thm:clt}
Assume the following mild regularity conditions:
\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
\item the model is correctly specified, identifiable, and consistent;
\item all data points in $Z_i$ are i.i.d.;
\item the parameter space $\Theta$ is open;
\item the matrix $\I_{\wtave}$ is non-singular;
\item the support set $A = \{(y,\x) : f(y,\trans\x\theta) > 0\}$ is independent of $\theta$;
and \item for all $(y,\x)\in A$, $f(y,\trans\x\theta)$ is three times differentiable with respect to $\theta$,
and the third derivative is continuous.
\end{enumerate}
Then in the limit as $n\to\infty$,
\begin{equation*}
\wtave - \wave_i \law \frac{1}{\sqrt n} \normal{0}{\I^{-1}_\wstar}
\end{equation*}
Here the notation $\law$ denotes convergence in law,
which is the standard notion of convergence for a sequence of probability distributions.
\end{theorem}

Note that Theorem \ref{thm:clt} does not require the likelihood $f$ to be convex.

There are many generalizations of Theorem \ref{thm:clt}.
For example, Theorem 5.2 in \cite{spokoiny2012parametricestimation} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
Because Spokoiny's theorem is non-asymptotic,
the distribution of the estimator is a sub-Gaussian random variable.
This more general setting requires more extensive notation,
but all of our results hold.

\subsection {Proof of Theorem 1}

We have by the triangle inequality that
\begin{equation}
\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wave} + \ltwo{\E\wave-\wave}
\label{eq:biasvar}
\end{equation}
The left term above captures the estimator's bias and the right term the variance.
First we consider the bias.
By the linearity of expectation, we have that
\begin{align}
\E\wave
&=
\E\frac{1}{m}\sum_{i=1}^m\wmle_i
%\\&=
=
\frac{1}{m}\sum_{i=1}^m\E\wmle_i
=
\E\wmle_i
\label{eq:expwave}
\end{align}
and so the bias term
$\ltwo{\wstar-\E\wave}
=
\ltwo{\wstar-\E\wmle_i}
$.
In words, the bias of the combined averaging estimator is the same as the bias of the estimator on an individual machine.

>>>>

%\newcommand{\smin}[1]{s_\text{min}({#1})}
\newcommand{\smin}{s_\text{min}}
\newcommand{\smax}{s_\text{max}}
\newcommand{\zero}{\text{\textbf{0}}}
\newcommand{\Zproj}{Z^{\textit{proj}}}
\newcommand{\Zreopt}{Z^{\textit{owa}}}
\newcommand{\nreopt}{n^{\textit{owa}}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\matW}{\hat W}
\newcommand{\matV}{\hat V}
\newcommand{\W}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Wowa}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Waff}{\mathcal{\hat A}}
\newcommand{\WaffE}{{\mathcal{\hat A}_\E}}
\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\ahat}{\hat\alpha}
\newcommand{\afull}{\ahat^{\textit{full}}}
\newcommand{\wowa}{\hat\w^{owa}}
\newcommand{\wowafull}{\hat\w^{\textit{owa,full}}}
\newcommand{\wowastar}{\hat\w^{\textit{owa,*}}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\wtave}{\E\hat\w^{ave}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

@@

@article{scikit-learn,
 title={Scikit-learn: Machine Learning in {P}ython},
 author={Pedregosa, F. and Varoquaux, G. and Gramfort, A. and Michel, V.
         and Thirion, B. and Grisel, O. and Blondel, M. and Prettenhofer, P.
         and Weiss, R. and Dubourg, V. and Vanderplas, J. and Passos, A. and
         Cournapeau, D. and Brucher, M. and Perrot, M. and Duchesnay, E.},
 journal={Journal of Machine Learning Research},
 volume={12},
 pages={2825--2830},
 year={2011}
}

@article{kddcup2012,
  author={Yanzhi Niu and Yi Wang and Gordon Sun and Aden Yue and Brian Dalessandro and Claudia Perlich and Ben Hamner},
  title = {The Tencent Dataset and KDD-Cup'12},
  howpublished = {URL \url{http://www.kddcup2012.org/c/kddcup2012-track2}},
  year = {2012}
}

@@

%\begin{assumption}
%The Fisher information matrix $\I_\wstar$ is positive definite.
%\end{assumption}
%
%Condition 1 above implies that the inverse $\I_\wstar^{-1}$ and square root $\I_\wstar^{1/2}$ exist.

%That is, we assume there exists a constant $c$ such that
%\begin{equation}
%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{\frac{d}{n}}} \le \exp\left(-ct^2\right)
%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{d/n}} \le \exp\left(-ct^2\right)
%%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(-\sigma^2t^2/2\right)
%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(\frac{-\sigma^2t^2}{2}\right)
%.
%\end{equation}
\begin{assumption}
Let $t>0$ and $\sigma^2$ be a constant called the \emph{variance proxy}.
Then with probability at least $1-\exp(-\sigma^2t^2/2)$,
\begin{equation}
%\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
\sqrt n \ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < t
,
\end{equation}
where $\I_\wstar$ is the positive definite Fisher information matrix.
\end{assumption}

This is a mild condition that is known to hold in many situations of interest.
In the asymptotic regime when $n\to\infty$,
very strong results are known.
%this condition is known as root-$n$ consistency.
%In this case,
The variance proxy $\sigma^2=1$,
and the distribution of the estimation error is truly Gaussian.
Theorem 7.5.2 of \cite{lehmann1999elements} provides a particularly simple exposition.
Lehman's theorem requires only that $f$ be three times differentiable,
the data points be i.i.d.,
and some mild identifiability conditions.
More sophisticated analyses show that Condition 1 holds very generally.
For example, the i.i.d. assumption is not strictly necessary.
See for example \cite{spokoiny2012parametricestimation}.

Similar results hold in the non-asymptotic case $n<\infty$,
but known results are less strong.
Most non-asymptotic results of this form require that the data points be sub-Gaussian.
For example, \cite{negahban2009unified} considers the case when the data points are sub-Gaussian, the likelihood satisfies a ``restricted strong convexity condition,'' and the regularizer is decomposable.
More recently, \cite{sivakumar2015beyond} showed Condition 1 holds when the data are only sub-exponential.
The strongest non-asymptotic results on the estimation error known to the authors is due to \cite{spokoiny2012parametricestimation}.
Spokoiny does not require any conditions on the data,
but can show that Condition 1 is satisfied only up to $t < O(n)$.

Our condition is more general than the conditions required in previous work.
\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded.
This condition implies Condition 1 because every bounded random variable is automatically sub-Gaussian by definition.

In conclusion,
there are many ways to verify that Condition 1 holds.
Pick whichever one suits your problem.
%With this condition assumed, our results are elementary.

@@

\begin{theorem}
\label{thm:wave}
Assume Condition 1 holds.
Let $t>0$.
Then with probability at least $1-\exp(-t)$,
\begin{equation}
\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wmle_i} + \sqrt\frac{v_t}{mn}
\end{equation}
where
\begin{equation}
v_t =
\sigma^2
\left(
\tr{\I^{-1}_{\wstar}}
+ 2\sqrt{\tr \left({\I^{-2}_{\wstar}}\right)t}
+ 2\ltwo{\I^{-1}_{\wstar}}t
\right)
\label{eq:vt}
\end{equation}
\end{theorem}

@@

Now we consider the variance.
We have that
\begin{align}
\ltwobig{\wave-\E\wave}
&=
%\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave
%\\&=
%\ltwobig{\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave}
%\label{eq:var1}
%\\&=
\ltwobig{\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)}
\\&=
%\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
%\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\sqrt{n}\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\Delta_i}
%\\&\sim
%\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\subnormal{\sigma^2}}
%\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\subnormal{\sigma^2}}
%\\&\sim
%\frac{1}{\sqrt {mn}}\ltwobig{\I^{-1/2}_\wstar\subnormal{\sigma^2}}
%\frac{1}{\sqrt{mn}}\ltwobig{\I^{-1/2}_\wstar\sqrt{n}\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
%\\&=
%\frac{1}{\sqrt {nm}}\normal\zero{\I^{-1}_{\E\wave}}
\label{eq:var1}
%\\&\sim
%\frac{1}{m}\normal{m\E\wmle_i}{\frac{m}{n}I^{-1}}
%\\&\sim
%\normal{\E\wmle_i}{\frac{1}{nm}I^{-1}}
%\\
%\wave-\E\wmle_i
%&\sim
%\normal{\zero}{\frac{1}{nm}I^{-1}}
%\\
%\wave-\E\wave
%&\sim
\end{align}
where
\begin{equation}
\Delta_i = \sqrt n \I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)
.
\end{equation}
Notice that each of the $\Delta_i$ are i.i.d. centered sub-Gaussian random vectors.
Centered sub-Gaussian random vectors inherit the following basic property from centered Gaussian random vectors:
The sum of $m$ i.i.d. sub-Gaussians has the same distribution as $\sqrt{m}$ times the sub-Gaussian.
That is, $\sum_{i=1}^m \Delta_i \sim \sqrt{m}\Delta_1$.
Substituting into Equation \ref{eq:var1} gives
\begin{equation}
\ltwobig{\wave-\E\wave}
\sim
\frac{1}{\sqrt{mn}}\ltwobig{\I^{-1/2}_\wstar\Delta_1}
.
\label{eq:var2}
\end{equation}

%Centering the distribution and applying Equation \ref{eq:expwave} gives
%\begin{equation}
%\E\wave-\wave
%\sim
%\normal{\zero}{\frac{1}{nm}I^{-1}}
%\end{equation}
%In the last step we used the fact that adding a sub-Gaussian random variable to itself $m$ times is equivalent lto multiplying the Gaussian by $\sqrt{m}$.
%So bounding the variance reduces to bounding the norm of a Gaussian.

Our last step is to bound the norm in the right hand side of Equation \ref{eq:var2}.
Theorem 2.1 from \cite{hsu2012tail} gives a bound on the norm of a positive semidefinite matrix times a sub-Gaussian random variable.
Applying the theorem gives that %with probability at least $1-\exp(-t)$,
\begin{equation}
\prob{
    \ltwobig{\I^{-1/2}_\wstar\Delta_1} < \sqrt{v_t}
}
\ge 1-\exp(-t)
,
\end{equation}
where $v_t$ is defined as in Equation \ref{eq:vt}.
And so the variance of $\wave$ satisfies
\begin{equation}
\prob{
\ltwobig{\wave-\E\wave}
<
\sqrt\frac{v_t}{{mn}}
}
\ge 1-\exp(-t)
.
\label{eq:ct1}
\end{equation}
%and so
%\begin{equation}
%\ltwo{\E\wave-\wave} \le \sqrt\frac{v_t}{nm}
%\label{eq:ct1}
%\end{equation}
%where
%\begin{equation}
%\begin{aligned}
%v_t'
   %= \tr\left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)
%+ 2&\sqrt{\tr \left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)^2t}&
%\\+ 2\ltwo{\frac{1}{nm}\I^{-1}_{\wtave}}t&
%\end{aligned}
%\end{equation}
%Then by the linearity of trace and scalability of norms,
%\begin{equation}
%v_t' =
%\frac{1}{mn}\left(
%\tr\I^{-1}_{\wtave}
%+ 2\sqrt{\tr \I^{-2}_{\wtave}t}
%+ 2\ltwo{\I^{-1}_{\wtave}}t
%\right)
%\label{eq:ct2}
%\end{equation}
Substituting Equations \ref{eq:expwave} and \ref{eq:ct1} into Equation \ref{eq:biasvar} gives the stated result.
\end{proof}

Note that this reduction in variance is essentially optimal.
If we apply Condition 2 to the estimator $\wmle$ instead of $\wmle_i$,
then we are assuming that the variance of $\wmle$ reduces at this same rate.
%By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
%we can show that the variance of the maximum likelihood estimator trained on all of the data is
%\begin{equation}
%\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
%\label{eq:var3}
%\end{equation}
%The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.

@@

\begin{lemma}
Assume Condition 1 holds.
Let $t>0$.
Then we have that with probability at least $1-\exp((d-m)(-t+\ln(t+1)))$,
\begin{equation}
\ltwo{\wstar-\proj\W\wstar}
=
(t+1)\sqrt{\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
\end{equation}
where $\smin$ and $\smax$ are the minimum and maximum eigenvalues of $\I_\wstar$.
\end{lemma}

\begin{proof}
We have that
\begin{align}
%&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
\ltwo{\wstar-\proj\W\wstar}
%\nonumber
%\\
&=
\ltwo{(\wstar-\wave) - (\proj\W\wstar - \wave)}
\nonumber
\\
&=
\ltwo{(\wstar-\wave) - \proj\W(\wstar - \wave)}
\nonumber
\\
&=
\ltwo{(I-\proj\W)(\wstar - \wave)}
.
\label{eq:i-projW}
\end{align}
Recall that we defined $\matW$ to be the $d\times m$ matrix with $i$th column equal $\wmle_i$.
The column space of $\matW$ is $\W$.
So the projection $\proj\W = \matW\matW^+$, where the superscript $^+$ is the Moore-Penrose pseudoinverse.
Define the matrix $\matV=\sqrt{n}I^{1/2}\matW$.
Then the columns of $\matV$ are sub-Gaussian by Condition 1,
and we can write
\begin{equation}
\proj\W
= \I^{-1/2} \matV \matV ^+ \I^{1/2}
.
\label{eq:vvp}
\end{equation}
Let $U\Sigma \trans V$ be the singular value decomposition of $\matV$,
where $U$ and $V$ are orthogonal matrices and $\Sigma$ is the diagonal matrix of singular values.
By the rotational invariance of the sub-Gaussian distribution,
$U$ and $V$ are distributed uniformly over the orthogonal group of matrices.
Substituting into Equation \ref{eq:vvp} gives
\begin{align}
\proj\W
&= \I^{-1/2} (U\Sigma V^T)(U \Sigma V^T) ^+ \I^{1/2}
\\
&= \I^{-1/2} (U\Sigma V^T)(V \Sigma^+ U^T) \I^{1/2}
\\
&= \I^{-1/2} U\Sigma \Sigma^+ U^T \I^{1/2}
.
\label{eq:vvp}
\end{align}
Substituting into Equation \ref{eq:i-projW} gives
\begin{align}
~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
\nonumber
\\
&=
\ltwo{(I-\I^{-1/2} U\Sigma \Sigma^+ U^T \I^{1/2})(\wstar - \wave)}
\\
&=
\ltwo{\I^{-1/2} U(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
\\
&=
\sqrt{\smin}\ltwo{U(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
\\
&=
\sqrt{\smin}\ltwo{(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
\end{align}
%The last line above follows from the rotational invariance of the $L2$ norm.
The operator $(I-\Sigma\Sigma^+)\trans U$ projects a vector onto a randomly chosen subspace of dimension $d-m$.
Lemma 2.2 from \cite{dasgupta2003elementary} provides a bound for this situation.
For all $t>0$, with probability at least $1-\exp((d-m)(-t+\ln (t+1)))$,
\begin{align}
\ltwo{(I-\Sigma\Sigma^+)\trans U \x} \le (t+1) \sqrt{1-\frac{m}{d}}
.
\end{align}
Substituting gives
\begin{align}
~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
\nonumber
\\
&=
(t+1)\sqrt{\smin\left(1-\frac{m}{d}\right)}\ltwo{\I^{1/2}(\wstar - \wave)}
\\
&=
(t+1)\sqrt{\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
\end{align}

\ignore{
\begin{theorem}
Assume Condition 1 holds.
Let $t>0$.
Then we have that with probability at least $(1-\exp(-t))^2$,
\begin{equation}
\ltwo{\wstar-\proj\W\wstar}
\le
%b_t\sqrt{\left(1-\frac{m}{d}\right)}
b_t\sqrt{\left(1-\frac{m}{d}\right)}
+
\sqrt\frac{v_{t}}{nm}
\end{equation}
where
%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
\begin{equation}
b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
\end{equation}
%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
and $v_t$ is defined as in Theorem \ref{thm:wave}.
\end{theorem}

This proof follows the same form as the proof for Theorem \ref{thm:wave}.
We first decompose the error into a bias and variance component;
then we bound each component separately.

Before we begin, we need to introduce the following vector space:
\begin{align}
\V &= \vecspan \{\wmle_i - \wtave\}_{i=1}^m
%,\\
%A &= \{ \vv + \E\wave : \vv \in \V \}
%,\\
%A &= \{ \vv + \wave : \vv \in \V \}
%.
\end{align}
%We also introduce notation for constructing affine spaces:
%\begin{equation}
%\V + \x = \{ \vv + \x : \vv \in \V \}
%.
%\end{equation}
%We now have that
We now have that:
\begin{align}
~~~~&\!\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
%&\ltwo{\wstar-\proj\W\wstar}
\nonumber
\\
&\le
\ltwo{(\wstar-\wtave) - \proj\W(\wstar-\wtave)}
\\
&\begin{aligned}
&\le
%\ltwo{
\lVert(\wstar-\wtave)
\\
&~~~
- (\proj\V(\wstar-\wtave) + \wtave - \wave)
%}
\rVert
\end{aligned}
\\
&\begin{aligned}
&\le
\ltwo{(\wstar-\wtave) - \proj\V(\wstar-\wtave)}
\\
&~~~
+ \ltwo{\wtave - \wave}
\end{aligned}
\\
&=
\ltwo{(I-\proj\V)(\wstar-\wtave)} + \ltwo{\wtave - \wave}
%
%
%
%\\
%&\le
%\ltwo{\wstar-(\proj{\V}\wstar + \wave)}
%\label{eq:step1}
%\\
%&\!\begin{aligned}
%&\le
%\ltwo{\wstar-(\proj{\V}\wstar+\wtave)}
%\\&~~~
%+
%\ltwo{(\proj{\V}\wstar+\wtave)-(\proj{\V}\wstar+\wave)}
%\end{aligned}
%\label{eq:step2}
%\\
%&=
%\ltwo{\wstar-(\proj{\V}\wstar+\wtave)}
%+
%\ltwo{\wtave-\wave}
%\label{eq:step3}
\end{align}
Inequality \ref{eq:step1} follows because $(\proj\V\wstar + \wave) \in \W$.
Inequality \ref{eq:step2} follows from the triangle inequality.
The rightmost term in Equation \ref{eq:step3} is the variance of $\wave$,
which we previously bounded with high probability in Equation \ref{eq:ct1}.

The affine space $\V+\wave$ is a subset of $\W$.
So,
\begin{align}
\ltwo{\wstar-\proj\W\wstar}
&\le
\ltwo{\wstar-\proj{(\V+\wave)}\wstar}
\end{align}
By the triangle inequality, we have
\begin{align}
%\\
&\begin{aligned}
\ltwo{\wstar-\proj\W\wstar}
&\le
\ltwo{\wstar-\proj{(\V+\E\wave)}\wstar}
\\&~~~~
+
\ltwo{\proj{(\V+\E\wave)}\wstar-\proj{(\V+\wave)}\wstar}
\end{aligned}
%\\&+
%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
\end{align}
The left term of the inequality above can be thought of as akin to the bias of the ``estimator'' $\proj\Waff\wstar$,
and the right term can be thought of as akin to the variance.

We begin by bounding the right term.
We have that with probability at least $1-\exp(-t)$,
\begin{align}
&\ltwo{\proj{(\V+\wtave)}\wstar-\proj{(\V+\wave)}\wstar}
\\
&=
\ltwo{(\proj{(\V+\wave)}\wstar + \wave - \wtave) -\proj{(\V+\wave)}\wstar}
%\ltwo{(\proj\Waff\wstar+\wave-\wtave)-\proj\Waff\wstar}
\\&=
\ltwo{\wave-\wtave}
\\&\le
\sqrt\frac{v_t}{mn}
\end{align}
%The first equality follows because the space $\WaffE$ is a translation of $\Waff$ by the vector $(\wave-\wtave)$.
The probabilistic inequality is due to Equation \ref{eq:ct1} in the proof of Theorem \ref{thm:wave}.
This shows that the variance of our estimator is the same as the variance of the averaging estimator,
which we already argued is essentially optimal.
